---
title: ggplot2bdc README
author: Brian Connelly
---

# ggplot2bdc

ggplot2bdc is a collection of themes and other things that can be used when
creating plots with [ggplot2](http://ggplot2.org/).

# Installation

`ggplot2bdc` is not on [CRAN](http://cran.r-project.org/), but you can use
[devtools](http://cran.r-project.org/web/packages/devtools/index.html) to
install the latest and greatest version. To do so:

    library(devtools)
    install_github('briandconnelly/ggplot2bdc')


# Examples

To gain access to the following themes, etc. provided, first load the
`gplot2bdc` package

```{r}
library(ggplot2bdc)
```

## theme_bdc_simplefacets

`theme_bdc_simplefacets` is a clean theme that uses thin black panels on a white
background. Legends are placed above. Facets labels are displayed as black text.
Subtle grid lines can be added to the panels with the `grid.x` and `grid.y`
parameters.

```{r theme_bdc_simplefacets,warning=FALSE}
ggplot(movies, aes(x=year, y=budget/1000000, color=rating)) +
    geom_point() +
    facet_grid(mpaa ~ .) +
    labs(x='Year', y='Budget ($ Millions)', title='Hollywood Budgets Against Time') +
    theme_bdc_simplefacets(grid.y=TRUE, grid.x=FALSE)
```


## theme_bdc_grey

`theme_bdc_grey` is a clean theme that uses thin grey panels on a white
background. Legends are placed above. Facet labels are displayed as black text
on a grey background.

```{r theme_bdc_grey,warning=FALSE}
ggplot(movies, aes(x=year, y=budget/1000000, color=rating)) +
    geom_point() +
    facet_grid(mpaa ~ .) +
    labs(x='Year', y='Budget ($ Millions)', title='Hollywood Budgets Against Time') +
    theme_bdc_grey()
```

Subtle grid lines can be added to the panels with the `grid.x` and `grid.y`
parameters:

```{r theme_bdc_grey grid,warning=FALSE}
ggplot(movies, aes(x=year, y=budget/1000000, color=rating)) +
    geom_point() +
    facet_grid(mpaa ~ .) +
    labs(x='Year', y='Budget ($ Millions)', title='Hollywood Budgets Against Time') +
    theme_bdc_grey(grid.x=FALSE, grid.y=TRUE)
```

## theme_bdc_black

`theme_bdc_black` is a clean theme that uses thin black panels on a white
background. Legends are placed above. Facet labels are displayed as black text
on a grey background. Subtle grid lines can be added to the panels with the
`grid.x` and `grid.y` parameters.

```{r theme_bdc_black,warning=FALSE}
ggplot(movies, aes(x=year, y=budget/1000000, color=rating)) +
    geom_point() +
    facet_grid(mpaa ~ .) +
    labs(x='Year', y='Budget ($ Millions)', title='Hollywood Budgets Against Time') +
    theme_bdc_black(grid.y=TRUE, grid.x=FALSE)
```


## theme_bdc_simple

`theme_bdc_simple` is a classic theme that displays plot axes as solid black
lines on a white background. This theme is probably best suited to single-panel
plots.

```{r theme_bdc_simple,warning=FALSE}
ggplot(movies, aes(x=year, y=length, color=rating, shape=mpaa)) +
    geom_point() +
    labs(x='Year', y='Length (Minutes)', title='Movie Lengths Against Time') +
    scale_y_log10() +
    theme_bdc_simple()
```

## coord_golden

`coord_golden` sets the aspect ratio of the axes according to the golden ratio.
For an introduction to how aspect ratio affects perception, see
[Aspect Ratio and Banking to 45 Degrees](http://eagereyes.org/basics/banking-45-degrees).

```{r coord_golden,warning=FALSE}
library(dplyr)

pg_movies <- filter(movies, mpaa=='PG')

ggplot(pg_movies, aes(x=year, y=budget/10^6, color=rating)) +
    geom_point() +
    labs(x='Year', y='Budget ($ Millions)', title='PG-13 Movie Budgets') +
    theme_bdc_grey() +
    coord_golden(xvals=pg_movies$year, yvals=pg_movies$budget/10^6)
```

## coord_square

`coord_square` sets the aspect ratio of the axes according to a square.

```{r coord_square,warning=FALSE,message=FALSE,error=FALSE}
library(dplyr)

pg_movies <- filter(movies, mpaa=='PG')

ggplot(pg_movies, aes(x=budget/10^6, y=rating)) +
    geom_smooth() +
    geom_point(shape=1) +
    labs(x='Budget ($ Millions)', y='Rating', title='Does Budget Affect Movie Ratings?') +
    theme_bdc_grey() +
    coord_square(xvals=pg_movies$budget/10^6, yvals=movies$rating)
```
