---
title: ggplot2bdc README
author: Brian Connelly
output: md_document
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.path="figure/", warning=FALSE, message=FALSE,
                      error=FALSE)
```

# ggplot2bdc

ggplot2bdc is a collection of themes and other things that can be used when
creating plots with [ggplot2](http://ggplot2.org/).


## Installation

`ggplot2bdc` is not on [CRAN](http://cran.r-project.org/), but you can use
[devtools](http://cran.r-project.org/web/packages/devtools/index.html) to
install the latest and greatest version. To do so:

```{r Installation, eval=FALSE}
if(!require("devtools")) install.packages("devtools")
devtools::install_github("briandconnelly/ggplot2bdc")
```

## Examples

To gain access to the following themes, etc. provided, first load the
`gplot2bdc` package. We'll also load `ggplot`, `dplyr`, and `magrittr` for
these examples.

```{r}
library(ggplot2bdc)
library(ggplot2)
library(dplyr)
library(magrittr)
```


### theme_bdc_grey

`theme_bdc_grey` is a clean theme that uses thin grey panels on a white
background. Legends are placed above. Facet labels are displayed as black text
on a grey background.

```{r theme_bdc_grey}
ggplot(data = economics, aes(x = date, y = unemploy)) +
    geom_line() +
    scale_x_date() +
    labs(x = "Year", y = "Thousands of Persons",
         title = "Monthly Unemployment", subtitle = "Seasonally Adjusted",
         caption = "Source: Federal Reserve Bank of St. Louis") +
    theme_bdc_grey()
```

Subtle grid lines can be added to the panels with the `grid.x` and `grid.y`
parameters.

```{r theme_bdc_grey_grid}
midwest$inmetro <- factor(midwest$inmetro, levels = c(0, 1),
                          labels = c("Rural", "Metro"))

ggplot(data = midwest,
       aes(x = inmetro, y = percbelowpoverty, color = state, shape = state)) +
    stat_summary(fun.y = "mean", geom = "line", aes(group = state)) +
    stat_summary(fun.y = "mean", geom = "point") +
    scale_color_hue(name = "State") +
    scale_shape_discrete(name = "State") +
    labs(x = NULL, y = "Percent of Population",
         title = "Poverty in the Midwest") +
    coord_equal(ratio = 0.5) +
    theme_bdc_grey(ticks.x = FALSE, grid.y = TRUE)
```

Some people prefer to display categorical data without tick marks along the
axis. These are included by default, but can be removed by setting either the
`ticks.x` or `ticks.y` parameters to `FALSE`.


### theme_bdc_simplefacets

`theme_bdc_simplefacets` is a modification of `theme_bdc_grey` where facet
labels are displayed as grey text on a white background.

```{r theme_bdc_simplefacets}
txcities <- txhousing %>%
    filter(city %in% c("Austin", "Dallas", "El Paso", "Fort Worth", "Houston",
                       "San Antonio"))

ggplot(data = txcities, aes(x = year, y = median / 1000)) +
    facet_grid(city ~ .) +
    stat_summary(fun.data = "mean_cl_boot", geom = "ribbon",
                 color = NA, alpha = 0.3) +
    stat_summary(fun.y = "mean", geom = "line") +
    labs(x = "Year", y = "Thousands of Dollars",
         title = "Home Prices on the Rise",
         subtitle = "Median Home Prices in Texas' Six Largest Cities",
         caption = "Source: TAMU Real Estate Center") +
    theme_bdc_simplefacets()
```


### theme_bdc_paneled

`theme_bdc_paneled` is a clean theme that displays plots without borders on grey
panels. Legends are placed above. Facet labels are displayed as plan black text.
Subtle grid lines can be added to the panels with the `grid.x` and `grid.y`
parameters, and tick markscan be managed with `ticks.x` and `ticks.y`. By
default, grid lines are used along the Y axis. This theme is best suited for
plots with multiple facets.

```{r theme_bdc_paneled}
ggplot(data = txcities, aes(x = year, y = median / 1000)) +
    facet_grid(city ~ .) +
    stat_summary(fun.data = "mean_cl_boot", geom = "ribbon",
                 color = NA, alpha = 0.3) +
    stat_summary(fun.y = "mean", geom = "line") +
    labs(x = "Year", y = "Thousands of Dollars",
         title = "Home Prices are Getting Bigger in Texas",
         subtitle = "Median Home Prices in Texas' Six Largest Cities") +
    theme_bdc_paneled()
```

To more clearly distinguish the individual panels, thick lines can be added on
the far right of each panel. For example:

```{r theme_bdc_paneled2, eval=FALSE}
# TODO: update me
ggplot(movies, aes(x = year, y = budget / 1000000, color = rating)) +
    geom_point() +
    facet_grid(mpaa ~ .) +
    labs(x="Year", y="Budget ($ Millions)",
         title="Hollywood Budgets Against Time") +
    geom_vline(aes(xintercept = Inf), color = "grey40", size = 3) +
    theme_bdc_paneled()
```


### theme_bdc_microtiter

`theme_bdc_microtiter` is a specialized theme for use in creating figures that
represent 96-well microtiter plates.

```{r loadplatemap,echo=FALSE}
library(dplyr)
platemap <- read.csv("~/Dropbox/Research/Documents/Blog/Blog Post-Analyzing Bacterial Growth/data/platemap.csv")
platemap <- mutate(platemap, Row=as.numeric(match(toupper(substr(Well,1,1)), LETTERS)), Column=as.numeric(substr(Well,2,5)))
platemap$Environment <- as.factor(platemap$Environment)
```

```{r theme_bdc_microtiter}
ggplot(data = platemap, aes(x = Column, y = Row)) +
    geom_point(data = expand.grid(Column = seq(1,12), Row = seq(1,8)),
               color = "grey90", fill = "white", shape = 21, size = 8) +
    geom_point(aes(shape = Environment, color = Strain), size = 9) +
    coord_fixed(ratio = (13/12)/(9/8), xlim = c(0.5, 12.5), ylim = c(0.6, 8.4)) +
    scale_y_reverse(breaks = seq(1,8), labels = LETTERS[1:8]) +
    scale_x_continuous(breaks = seq(1,12)) +
    labs(title = "Plate Layout for My Experiment", subtitle = "25 March 2016") +
    theme_bdc_microtiter()
```

For more information, see the [Plotting Microtiter Plate Maps](http://bconnelly.net/2014/05/plotting-microtiter-plate-maps/).

### theme_bdc_lattice_population

`theme_bdc_lattice_population` is a specialized theme intended to display
populations of individuals structured spatially in a lattice.

```{r theme_bdc_lattice_population}
ggplot(data = filter(rockpaperscissors, Time == 10),
       aes(x = X, y = Y, color = Type)) +
        geom_point(shape=15) +
        coord_equal(ratio=1) +
        scale_color_hue() +
        theme_bdc_lattice_population()
```

We can also show multiple population states:

```{r theme_bdc_lattice_population-facets}
ggplot(data = filter(rockpaperscissors, Time %in% c(0, 200, 400, 600)),
       aes(x = X, y = Y, color = Type)) +
    facet_wrap(~Time, nrow = 1, ncol = 4) +
    geom_point(shape = 15) +
    coord_equal(ratio = 1) +
    scale_color_hue(guide = FALSE) +
    labs(title = "Rock, Paper, Scissors",
         subtitle = "Wait, is it ON three or AFTER three?",
         caption = "Data from simulation on a 50x50 lattice") +
    theme_bdc_lattice_population()
```

### rescale_golden

`rescale_golden` adjusts the proportions of the axes of the given plot object
so that follow the golden ratio (horizontally). If no plot is provided, the last
plot that was displayed is used.

```{r rescale_golden}
ggplot(mtcars, aes(x = hp, y = mpg)) +
    geom_smooth() +
    geom_point(shape = 1) +
    labs(x = "Horsepower", y = "Fuel Efficiency (mpg)") +
    theme_bdc_grey()
```

```{r rescale_golden2}
rescale_golden()
```


### rescale_square

`rescale_square` adjusts the proportions of the axes of the given plot object
so that they are equal. If no plot is provided, the last plot that was displayed
is used.

```{r rescale_square}
ggplot(data = economics, aes(x = date, y = unemploy)) +
    geom_line() +
    labs(x = "Date", y = "Thousands of Unemployed Persons") +
    theme_bdc_grey()
```

```{r rescale_square2}
rescale_square()
```


### rescale_plot

`rescale_plot` adjusts the proportions of the axes of the given plot object to
the given ratio. If no plot is provided, the last plot that was displayed is
used.

```{r rescale_plot}
ggplot(data = economics, aes(x = date, y = unemploy)) +
    geom_line() +
    labs(x = "Date", y = "Unemployed Persons (x1000)") +
    theme_bdc_grey()
rescale_plot(ratio = 16 / 9)
```

### coord_golden

`coord_golden` sets the aspect ratio of the axes according to the golden ratio.
Whether the panel follows the golden ratio horizontally or vertically can be set
using the `orientation` argument. For an introduction to how aspect ratio
affects perception, see
[Aspect Ratio and Banking to 45 Degrees](http://eagereyes.org/basics/banking-45-degrees).

```{r coord_golden, eval=FALSE}
# TODO: update this
pg_movies <- filter(movies, mpaa == "PG")

ggplot(pg_movies, aes(x=year, y=budget / 10^6, color = rating)) +
    geom_point() +
    labs(x="Year", y="Budget ($ Millions)", title="PG-13 Movie Budgets") +
    theme_bdc_grey() +
    coord_golden(xvals=pg_movies$year, yvals=pg_movies$budget / 10^6,
                 orientation="horizontal")
```

### coord_square

`coord_square` sets the aspect ratio of the axes according to a square.

```{r coord_square, eval=FALSE}
# TODO update this
pg_movies <- filter(movies, mpaa == "PG")

ggplot(pg_movies, aes(x = budget / 10^6, y = rating)) +
    geom_smooth() +
    geom_point(shape = 1) +
    labs(x="Budget ($ Millions)", y="Rating",
         title="Does Budget Affect Movie Ratings?") +
    theme_bdc_grey() +
    coord_square(xvals = pg_movies$budget / 10^6, yvals = movies$rating)
```


### ggsave_golden

`ggsave_golden` saves the given plot to the given filename. If no plot is
specified, the most recently displayed plot is used. The proportions of the
plot's axes should first be set to the golden ratio using `rescale_golden` or
`coord_golden`.

```{r ggsave_golden}
p <- ggplot(mtcars, aes(x = hp, y = mpg)) +
    geom_smooth() +
    geom_point(shape = 1) +
    labs(x = "Horsepower", y = "Fuel Efficiency (mpg)") +
    theme_bdc_grey()
p <- rescale_golden(plot = p)
ggsave_golden(filename = "mpg_against_horsepower.pdf", plot = p)
```
